\documentclass[times, twoside]{zHenriquesLab-StyleBioRxiv}

\usepackage{newfloat}
\DeclareFloatingEnvironment[name={Supplementary Figure}]{suppfigure}
\DeclareFloatingEnvironment[name={Supplementary Table}]{supptable}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{upgreek}

\newcommand{\ecoli}{\textit{E. coli}}
\newcommand{\dreca}{$\Delta$\textit{recA}} % ΔrecA
\newcommand{\teneighty}{RecB$_{1080}$} % RecB1080
\newcommand{\geneteneighty}{\textit{recB$_{1080}$}} % recB1080
\newcommand{\od}{OD$_{600}$} % OD600
\newcommand{\umsq}{$\upmu$m$^2$} % µm2
\newcommand{\ums}{$\upmu$m$^2$.s$^{-1}$} % µm2.s-1
\newcommand{\ncells}[1]{N$_{\text{cells}}$ = #1} % N_cells = {}
\newcommand{\nspots}[1]{N$_{\text{spots}}$ = #1} % N_spots = {}
\newcommand{\nnucl}[1]{N$_{\text{nucleoids}}$ = #1} % N_nucleoids = {}

% Surname of the lead author for the running footer
\leadauthor{Thédié} 

\begin{document}

\title{Spatio-temporal coordination of the DNA Double Strand Break repair machinery in \emph{Escherichia coli}}
\shorttitle{Spatio-temporal dynamics of the DSB repair machinery in \ecoli}

\author[1]{Daniel Thédié}
\author[1,2]{Alessia Lepore}
\author[1]{Lorna McLaren}
\author[1,3]{Meriem El Karoui}


\affil[1]{Institute of Cell Biology, University of Edinburgh, Edinburgh, UK}
\affil[2]{Laboratory for Optics and Biosciences, \'Ecole Polytechnique, Institut Polytechnique de Paris, Palaiseau, France}
\affil[3]{Laboratoire de Biologie et Pharmacologie Appliquée, CNRS,Ecole Normale Supérieure Paris-Saclay, UMR 8113, Gif-sur-Yvette, France}  % ENS Saclay, to be checked by Meriem

\maketitle

\begin{abstract}
    Repairing DNA damage is of primary importance for all living organisms. DNA double-strand breaks (DSBs) are one of the most serious types of DNA damage, as they lead to loss of genetic information and death when not repaired. In \ecoli, they are recognised and processed by the RecBCD complex, which initiates repair by homologous recombination. Although the repair dynamics downstream of RecBCD has been well characterised, it is still unclear how long this complex stays attached to DNA and what triggers its dissociation \emph{in vivo}. To answer these questions, we imaged RecB at the single-molecule level, and quantified its dynamic behaviour in bacterial cells exposed to ciprofloxacin, an antibiotic that induces DSBs. Our results show that RecB forms long-lived complexes with DSBs (10 seconds), and that its dissociation from DNA is an intrinsic property of the complex, that does not depend on the amount of DNA damage, nor the following steps in the repair pathway. Moreover, we show that we can use RecB binding to DSBs as a marker to estimate the rate of damage formation. This study provides a detailed quantitative insight into the interaction of RecBCD with DNA double-strand ends in \ecoli\ \emph{in vivo}, and into the bacterial response to DSBs induced by ciprofloxacin.
\end{abstract}

\begin{keywords}
    DNA repair | RecBCD | Imaging | \emph{in vivo}
\end{keywords}

\begin{corrauthor}
    meriem.elkaroui\at ed.ac.uk
\end{corrauthor}

\input{Introduction}

\input{Methods}

\input{Results}

\input{Discussion}

\begin{funding}
    This work has been supported by Wellcome Trust Investigator Awards (Grant No. 205008/Z/16/Z awarded to M.E.K.), a BBSRC BB/S008012/1 responsive mode award (to M.E.K.), and a Marie Skłodowska-Curie Personal Fellowship (Grant No. 101063725-BARTAS) awarded to A.L.
\end{funding}

\begin{acknowledgements}
    We would like to thank Prof. David Leach for insightful discussions on DNA repair in \emph{E. coli}. We thank Prof. Mark Dillingham for sharing his knowledge on the biochemistry of RecBCD, and for gifting the Gam plasmid. We thank Prof. Johan Elf for kindly sharing the RecA-SYFP2 tandem fusion strain. We thank Dr. Jean Ollion for the development of BACMMAN and its associated tools, as well as the image analysis support he provided. We thank Elise Darmon for discussing results, and reviewing the manuscript.
\end{acknowledgements}

\begin{contributions}
    M.E.K. and D.T. conceived the experiments and designed the data analysis. D.T., A.L., and L.McL. built the strains. D.T. collected and analysed the data. M.E.K., D.T. and A.L. discussed the data. D.T. wrote the manuscript's first draft and created the figures. M.E.K., D.T. and A.L. revised the manuscript. L.McL., as the lab manager, oversaw order placements and maintained the lab organisation. M.E.K. provided funding. All authors read, edited and approved the final manuscript.
\end{contributions}

\begin{interests}
    The authors declare no competing financial interests.
\end{interests}

\section*{Bibliography}
\bibliography{Bibliography}

\onecolumn
\newpage

\begin{center}
    \huge
    \vspace*{9em}
    \textbf{Supplementary information}\\
    \vspace*{2em}
    \textbf{Single molecule imaging of RecB \emph{in vivo} reveals dynamics of DNA Double Strand Break repair in \emph{Escherichia coli}}\\
    \vspace*{2em}
    \Large
    Daniel Thédié, Alessia Lepore, Lorna McLaren and Meriem El Karoui
\end{center}

\newpage

\input{Supplementary}

\end{document}
